### 2014-05-13 research/paper/main/article.tex
modifications 2.28 13 may

Fluid-elastic galloping is one of the sub-areas of research in fluid structure interactions. This area has been of interest due to the vibrations crated by galloping on transmission lines and civil structures and leading them to failure. Therefore understanding this phenomenon in order to suppresses these vibrations was quite important. However, the search for alternate energy sources with minimal environmental impact has become an important area of research in the modern word. Therefore researchers are moving towards investigating the possibility of extracting useful energy from this vibrations rather than suppressing them. Therefore, it is quite important to understand the governing parameters and analyse the influence of them on the energy transfer from the fluid to the body, because this understanding will lead to develop better practical applications. Therefore, in this paper we focus on understanding the energy transfer from the fluid to the body and isolate the governing parameters influencing it.

@@

However, the QSS model equation when solved analytically using the sinusoidal solution method cannot predict the response for cases with low mass ratios. \citet{Joly2012} observed that finite element simulations show a sudden change in amplitude below a critical value of the mass ratio. The model equation defined in \citet{Parkinson1964} was modified to account for the vortex shedding and solved numerically to predict the reduced amplitude at low mass ratios to the point where galloping is no longer present. \citet{Barrero-Gil2010a} investigated the possibility of extracting power from vibrations caused by galloping using the quasi-steady state model. In the conclusions of that paper it was pointed out that in order to obtain a high power to area ratio, the mass-damping ($m^*\zeta$) parameter should be kept low. The same study investigated the influence of the characteristics of the $C_y$ curve on maximum power output.

Here, the modified QSS model developed by \citet{Joly2012} is integrated numerically for low Reynolds numbers. The focus is on the power extraction potential as a function of mechanical parameters (i.e. frequency of oscillation, damping factor and mass ratio). To this end, a series of previously mentioned mechanical parameters are tested at two different values of \reynoldsnumber: $\reynoldsnumber = 165$, a case that should remain laminar and essentially two-dimensional; $\reynoldsnumber = 22300$, a case where the flow is expected to be turbulent and three-dimensional. Both cases require the input of transverse force coefficients $C_y$ as a function of angle of attack $\theta$ for a fixed body. These data are provided from direct numerical simulations for the $\reynoldsnumber = 165$ case, while the data provided by \citet{Parkinson1964} are used for the $\reynoldsnumber = 22300$ case.

The structure of the paper is as follows. Section \ref{sec:theory} presents the modified QSS model, the method for the calculation of power output, and the parameters used. Section \ref{sec:results} presents the results, first of the fixed body tests at a range of $\theta$, then of the response characteristics predicted by the integration of the QSS model for both the high and low \reynoldsnumber\ cases. For the low \reynoldsnumber\ case, the results of the QSS model are compared to those of full direct numerical simulations of the fluid-structure interaction problem. Finally, section \ref{sec:conc} presents the conclusions that can be drawn from this work.
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
\section*{Nomenclature}
%\textbf{Nomenclature}

@@

>>>>

Fluid-elastic galloping is one of the sub-areas of research in fluid structure interactions. This area has been of interest due to the vibrations crated by galloping on transmission lines and civil structures and leading them to failure. Therefore understanding this phenomenon in order to suppresses these vibrations was quite important. However, the search for alternate energy sources with minimal environmental impact has become an important area of research in the modern word. Therefore researchers are moving towards investigating the possibility of extracting useful energy from this vibrations rather than suppressing them. Thus, it is quite important to understand the governing parameters and analyse the influence of them on the energy transfer from the fluid to the structure, because this understanding will lead to develop better practical applications. Hence, in this paper we focus on understanding the energy transfer from the fluid to the body and isolate the governing parameters influencing it.

@@

However, the QSS model equation when solved analytically using the sinusoidal solution method cannot predict the response for cases with low mass ratios. \citet{Joly2012} observed that finite element simulations show a sudden change in amplitude below a critical value of the mass ratio. The quasi-oscillator equation in \citet{Parkinson1964} was altered to account for the vortex shedding and solved numerically to predict the reduced amplitude at low mass ratios to the point where galloping is no longer present. \citet{Barrero-Gil2010a} investigated the possibility of extracting power from vibrations caused by galloping using the quasi-steady state model. In the conclusions of that paper it was pointed out that in order to obtain a high power to area ratio, the mass-damping ($m^*\zeta$) parameter should be kept low. The same study investigated the influence of the characteristics of the $C_y$ curve on maximum power output.

Here, the modified QSS model developed by \citet{Joly2012} is integrated numerically for low Reynolds numbers. The focus is on the power transfer from the fluid to the structure and the influence of mechanical parameters (i.e. frequency of oscillation, damping factor and mass ratio). To this end, a series of previously mentioned mechanical parameters are tested at two different values of \reynoldsnumber: $\reynoldsnumber = 165$, a case that should remain laminar and essentially two-dimensional; $\reynoldsnumber = 22300$, a case where the flow is expected to be turbulent and three-dimensional. Both cases require the input of transverse force coefficients $C_y$ as a function of angle of attack $\theta$ for a fixed body. These data are provided from direct numerical simulations for the $\reynoldsnumber = 165$ case, while the data provided by \citet{Parkinson1964} are used for the $\reynoldsnumber = 22300$ case.

The structure of the paper is as follows. Section \ref{sec:theory} presents the governing equations and the oscillator model used to obtain data, the method for the calculation of the power transferred from the fluid to the structure. The governing parameters namely, the combined mass-stiffness and the combined mass-damping obtained using linearised time scales of the oscillator model are also being introduced. Section \ref{sec:results} presents the results, first of the fixed body tests at a range of $\theta$, then of the response characteristics predicted by the integration of the QSS model for both the high and low \reynoldsnumber\ cases. For the low \reynoldsnumber\ case, the results of the QSS model are compared to those of full direct numerical simulations of the fluid-structure interaction problem. Finally, section \ref{sec:conc} presents the conclusions that can be drawn from this work.
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
\section*{Nomenclature}
%\textbf{Nomenclature}

@@

 % % % % % % % % % % % % % Time scales % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
 The natural time scales of the system can be found by solving for the eigenvalues of the linearized equation of motion, namely
 \begin{equation}
 \label{eqn:eom_linear}
 (m+m_a)\ddot{y}{+}c\dot{y}{+}ky{=}\frac{1}{2}\rho U^2 \mathcal{A} a_1\left(\frac{\dot{y}}{U}\right),
 \end{equation}
 which is simply the equation of motion presented in equation \ref{final_equation_motion} with the polynomial series for the lift force truncated at the linear term, and the forcing term representing vortex shedding removed.

 Combining the $\dot{y}$ terms and solving for eigenvalues gives
 \begin{equation}
   \label{eqn:eigs}
   \lambda_{1,2}= -\frac{1}{2}\frac{c-\frac{1}{2}\rho U\mathcal{A}a_1}{(m+m_a)}\pm\frac{1}{2}\sqrt{\left[\frac{c-\frac{1}{2}\rho U\mathcal{A}a_1}{(m+m_a)}\right]^2-4\frac{k}{(m+m_a)}}.
 \end{equation}

 If it is assumed that the spring is relatively weak, $k\rightarrow 0$, a single non-zero eigenvalue remains. This eigenvalue is
 \begin{equation}
   \label{eqn:eigs_nospring}
   \lambda=-\frac{c-\frac{1}{2}\rho U\mathcal{A}a_1}{(m+m_a)}.
 \end{equation}

 Further, if it is assumed that the mechanical damping is significantly weaker than the aerodynamic forces on the body, $c\rightarrow 0$ and
 \begin{equation}
   \label{eqn:eigs_nospring_nodamp}
   \lambda=\frac{\frac{1}{2}\rho U\mathcal{A}a_1}{(m+m_a)}.
 \end{equation}

 For reasonably heavy bodies, the impact of the added mass can also be neglected, to arrive at a definition of $\lambda$ as
 \begin{equation}
   \label{eqn:eigs_nospring_nodamp}
   \lambda=\frac{\frac{1}{2}\rho U\mathcal{A}a_1}{m}.
 \end{equation}

 In this form, $\lambda$ represents the inverse time scale of the motion of the body due to the negative damping effect of the long-time aerodynamic forces. In fact, the terms can be regrouped and $\lambda$ written as
 \begin{equation}
   \label{eqn:timescale}
   \lambda = \frac{a_1}{m^*}\frac{U}{D}
 \end{equation}

 Written this way, the important parameters that dictate this inverse time scale are clear. The rate of change in the aerodynamic force with respect to angle of attack when the body is at the equilibrium position, $\partial C_y/\partial \alpha$, is represented by $a_1$. The mass ratio is represented by $m^*$. The inverse advective time scale of the incoming flow is represented by the ratio $U/D$. Increasing $a_1$ would mean the force on the body would increase more rapidly with small changes in the angle of attack, $\theta$, or transverse velocity. Equation \ref{eqn:timescale} shows that such a change will increase the inverse time scale, or analogously decrease the response time of the body. Increasing the mass of the body, thereby increasing $m^*$, has the opposite effect. The inverse time scale is decreased, or as might be expected, a heavier body will take longer to respond.

 This timescale can then be used to non-dimensionalize the equation of motion, and to find the relevant dimensionless groups of the problem. If the non-dimensional time, $\tau$, is defined such that $\tau=t(a_1/m^*)(U/D)$, the equation of motion presented in equation \ref{final_equation_motion} can be non-dimensionalized as
 \begin{equation}
   \label{eqn:eom_nondim}
   \ddot{Y} + \frac{m^{*2}}{a_1^2}\frac{kD^2}{mU^2}Y = \left(\frac{1}{2} - \frac{m^*}{a_1}\frac{cD}{mU}\right)\dot{Y} + H.O.T.,
 \end{equation}

 where $H.O.T.$ respresents the higher order terms in $\dot{Y}$. The coefficients can be regrouped into combinations of non-dimensional groups, and rewritten as
 \begin{equation}
   \label{eqn:eom_nondim_regroup}
   \ddot{Y} + \frac{4\pi^{2}m^{*2}}{U^{*2}a_1^2}Y = \left(\frac{1}{2} - \frac{c^*m^*}{a_1}\right)\dot{Y} + H.O.T,
 \end{equation}

 where $c^*=cD/mU$ is a non-dimensional damping parameter.

 Equation \ref{eqn:eom_nondim_regroup} shows there are four non-dimensional parameters that play a role in setting the response of the system. These are the stiffness (represented by the reduced velocity $U^*$), the damping $c^*$, the mass ratio $m^*$, and the geometry, represented by the rate of change in the aerodynamic force with respect to angle of attack when the body is at the equilibrium position, $a_1$. The grouping of these parameters into two groups in equation \ref{eqn:eom_nondim_regroup} which arise by non-dimensionalising using the natural time scale of the galloping system, suggests there are two groups that dictate the response: $\Gamma_1 = 4\pi^2m^{*2}/U^{*2}a_1^2$ and $\Gamma_2 = c^*m^*/a_1$. For a given geometry and Reynolds number, $\Gamma_1$ can be thought of as a combined mass-stiffness, whereas $\Gamma_2$ can be thought of a a combined mass-damping parameter. As it is assumed that during galloping the stiffnes plays only a minor role, $\Gamma_2$ seems a likely parameter to collapse the data presented in figure \ref{fig:uncollapsed_data}. In fact, in the classic paper on galloping from \citet{Parkinson1964}, galloping data from wind tunnel tests is presented in terms of a parameter that can be shown to be the same as $\Gamma_2$.

 All of the quantities that make up $\Gamma_1$ and $\Gamma_2$ can, in theory, be known before an experiment is conducted. However, the quantity $a_1$ is a relatively difficult one to determine, requiring static body experiments or simulations. Here, the geometry is unchanged and results are only being compared at the same \reynoldsnumber. Hence, suitable parameters can be formed by multiplying $\Gamma_1$ and $\Gamma_2$ by $a_1^2$ and $a_1$ respectively, to arrive at a mass-stiffness parameter $\massstiff =  4\pi^2m^{*2}/U^{*2}$, and a mass-damping parameter defined as $\massdamp = c^*m^*$.
